---
title: "HW6"
author: "Zach Steffens"
date: "2026-03-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(lubridate)


load("C:\\Users\\zrsm9\\Downloads\\house_prices.rda")

df_names = ls()
house = NULL
for (nm in df_names) {
  if (is.data.frame(get(nm))) {
    house = get(nm)
    break
  }
}

names(house)

house = house %>%
  mutate(
    date = as.Date(date),
    year = year(date)   
  )

p1 = ggplot(house, aes(x = year, y = house_price_index)) +
  geom_line(color = "steelblue") +
  facet_wrap(~ state, scales = "free_y") +
  scale_x_continuous(
    breaks = c(1980, 2000, 2020),
    labels = c("'80", "'00", "'20")
  ) +
  labs(
    title = "House Price Index Over Time by State",
    x = "Year",
    y = "House Price Index"
  ) +
theme_minimal() +
  theme(
    strip.text = element_text(size = 4),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.title.x = element_text(size = 8),
    axis.title.y = element_text(size = 8),
    legend.text = element_text(size = 6),
    legend.title = element_text(size = 7)
  )

print(p1)

#(b) Reshape
house_reshaped = house %>%
  gather(
    key = "measure",
    value = "value",
    -state, -date, -year
  )

head(house_reshaped)

#(c) Two-line plot 
house_reshaped2 = house_reshaped %>%
  filter(measure %in% c("house_price_index", "unemploy_perc"))

p2 = ggplot(house_reshaped2, aes(x = year, y = value, color = measure)) +
  geom_line() +
  facet_wrap(~ state, scales = "free_y", ncol = 9) +
  theme_minimal() +
  theme(
    strip.text = element_text(size = 4),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.title.x = element_text(size = 8),
    axis.title.y = element_text(size = 8),
    legend.text = element_text(size = 6),
    legend.title = element_text(size = 7)
  ) +
  scale_x_continuous(
    breaks = c(1980, 2000, 2020),
    labels = c("'80", "'00", "'20")
  ) +
  scale_color_manual(
    values = c("house_price_index" = "steelblue",
               "unemploy_perc" = "firebrick"),
    labels = c("House Price Index", "Unemployment %")
  ) +
  labs(
    title = "House Price Index and Unemployment % Over Time",
    x = "Year",
    y = "Value",
    color = "Measure"
  ) 

print(p2)
```
Overlaying price index and unemployement percentage on the same y-axis with 
facet wrap is not a good approach because the measures have different scales/units. 
This can lead to misleading comparisons or difficulty seeing trends, especially
given the relatively low value of unemployment percentage. An approach could be to 
normalize the values so they are equally weighted and then plot them. 
